---
title: "Homework 6"
author: "Manali Rege-Colt"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
#### Assign the variable n_dims a single random integer between 3-10:

```{r}

n_dims <- sample (3:10, 1)
print(n_dims)

```

#### Create a vector of consecutive integers from 1 to n_dims^2: 

```{r}

cons_int <- c(1:n_dims^2)
print(cons_int)

```

#### Use the sample function to randomly reshuffle these values:

```{r}

samp <- sample(cons_int)
print(samp)

```

#### Create a square matrix with these elements:

```{r}

sq_matrix <- matrix(samp,nrow=4,ncol=4)

```

#### Print the matrix

```{r}

print(sq_matrix)

```

#### Find a function in r to transpose the matrix:

```{r}

trans_matrix <- t(sq_matrix)

```

#### Print again to see changes:

```{r}

print(trans_matrix)

```

#### Calculate the sum and the mean of the elements in the first row and the last row:

```{r}

sum(trans_matrix[1,])

mean(trans_matrix[1,])

sum(trans_matrix[4,])

mean(trans_matrix[4,])

```

#### Read about the eigen() function and use it on your matrix:

```{r}

eigen<- eigen(trans_matrix, symmetric=TRUE, 
              only.values=FALSE, 
              EISPACK=FALSE)

```

#### look carefully at the elements of $values and $vectors. What kind of numbers are these?
##### $values is a vector of values that are sorted in decreasing order.
##### $vectors is a matrix compiled of numbers between -1 and 1.


#### Dig in with the typeof() function to figure out their type:

```{r}

typeof(eigen$values)

typeof(eigen$vectors) 

```

# Problem 2

#### Create a list with the following named elements:
* my_matrix, which is a 4 x 4 matrix filled with random uniform values
* my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
* my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}

my_matrix <- matrix(data=runif(16),ncol=4,nrow=4)
print(my_matrix)

my_vec <- c(1:100)
my_vec <- sample(my_vec)

my_logical <- my_vec < 50
print(my_logical)

my_letters <- c(LETTERS[1:26])
my_letters <- sample(my_letters)
print(my_letters)

my_list<- list(my_matrix,my_logical,my_letters)
str(my_list)

```

#### Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector:

```{r}

new_list <- list(my_matrix[2,2],my_logical[2],my_letters[2])
print(new_list)

```

#### Use the typeof() function to confirm the underlying data types of each component in this list:

```{r}

typeof(my_matrix[2,2])
typeof(my_logical[2])
typeof(my_letters[2])

```

#### Combine the underlying elements from the new list into a single atomic vector with the c() function:

```{r}

new_vec <- c(my_matrix[2,2],my_logical[2],my_letters[2])
print(new_vec)

```

#### What is the data type of this vector?

```{r}

typeof(new_vec)

```

# Problem 3
#### Create a data frame with two variables (= columns) and 26 cases (= rows) and call the first variable my_unis and fill it with 26 random uniform values from 0 to 10: 

```{r}

my_unis <- runif(26,0,10)
print(my_unis)

```

#### Call the second variable my_letters and fill it with 26 capital letters in random order.

```{r}

my_letters <- sample(LETTERS[1:26])
my_letters

data <- data.frame(my_unis,my_letters)
print(data)
```

#### For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA:

```{r}

my_unis[sample(1:26,4)] <- NA
print(my_unis)

```

#### For the first variable, write a single line of R code to identify which rows have the missing values:

```{r}

which(!complete.cases(my_unis))

```

#### For the second variable, sort it in alphabetical order:

```{r}

sort(my_letters)

```

#### Calculate the column mean for the first variable:

```{r}

mean(my_unis,na.rm=TRUE)

```




